---
title: "Vax forecast - DTP3"
author: "Beth Evans"
date: "10/09/2021"
output: 
  html_document:
    code_folding: "show"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set-up

### Install packages
```{r, message = FALSE}
    library(tidyverse)
    library(rio)
    library(magrittr)
    library(timetk)
    library(tidymodels)
    library(lubridate)
    library(ggplot2)
    library(ggforce)
    library(urca)
    library(tseries)
    library(forecast)
    library(reshape2)
    library(dplyr)
    library(data.table)
    library(countrycode)
```

### Load data
Import data:
- Coverage data from WUENIC for 2000-2020 <br>
- Income group classification from World Bank <br>
- Gavi classification pre-compiled dataset from Gavi website country documents and Eligibility policy <br>
```{r}
    file_path_dtp3 <- here::here("data", "dtp3.csv")
    data_raw <- file_path_dtp3 %>%
      rio::import(header = TRUE) %>%
      tibble()
    data_raw
    
    file_path_wb <- here::here("data", "wb_ig.csv")
    income <- file_path_wb %>%
      rio::import(header = TRUE) %>%
      tibble()
    income
    
    file_path_gavi <- here::here("data", "gavi.csv")
    gavi <- file_path_gavi %>%
      rio::import(header = TRUE) %>%
      tibble()
    gavi
```

### Clean data
Steps taken:
- rename variables 
- reshape data to long vs. wide format
- add categorisation columns (income group, region, Gavi eligibility status)
- select relevant columns only 
```{r}
    ##Clean and reshape 2000-2019 coverage data
      data_raw %<>%
        dplyr::rename(region = unicef_region, iso_code = iso3)
      
      data_long <- data_raw %>%
        dplyr::select(-"2020") %>%
        pivot_longer(-(region:vaccine), names_to = "year", values_to = "coverage")
      names(data_long) <- tolower(names(data_long))
      data_long
    
    ##Clean 2020 coverage data (actuals)
      cov_2020 <- data_raw %>%
        dplyr::select(iso_code, country, '2020') %>%
        dplyr::rename(actuals_2020 = '2020') %>%
        mutate(actuals_2020 = actuals_2020 / 100)
    
    ##Join 'income' classification to dataset
      income %<>% 
        filter(datayear == "2020") %>%
        dplyr::rename(iso_code = wb_code) %>%
        dplyr::select(-c(datayear, country, fiscalyear, publicationyear))
      income
    
      data_long %<>%
        inner_join(income, by = "iso_code")
      data_long
      
    ##Join 'Gavi' classification to dataset
      gavi %<>% 
        filter(Year == "2020") %>%
        dplyr::rename(iso_code = "WB CC", country = "LB Country", gavi_binary = "Gavi Eligibility", gavi_status = "Gavi Co-financing") %>%
        dplyr::select(-c(country, Year, "Transition Year", MatchValue))
      gavi
    
      data_long %<>%
        left_join(gavi, by = "iso_code") %>%
        mutate(gavi_binary = replace_na(gavi_binary, "Never Gavi"), gavi_status = replace_na(gavi_status, "Never Gavi")) 
      
    ##Clean consolidated projection dataset
      x <- data_long %>%
        mutate(year = as.integer(year), coverage = coverage / 100) %>%
        dplyr::select(-vaccine) %>%
        dplyr::select(region, country, iso_code, income_group, gavi_status, year, coverage) %>%
        arrange(country, year)
      x

```

## Data filtering

### Filter data for complete data
We remove:
- entries where coverage is NA
- years prior to 2000
- countries without data for the last 20 consecutive years

```{r}
    x <- x %>%
      filter(year >= 2000, !is.na(coverage))
    
    sufficient_data <- x %>%
      group_by(iso_code) %>%
      dplyr::summarize(count = n()) %>%
      filter(count >= 20) 
    
    x %<>%
      inner_join(sufficient_data, by = "iso_code") %>%
      dplyr::select(-count) %>%
      mutate(year = as.Date(ISOdate(as.numeric(year), 12, 31)))
    x
```

## Coverage modelling
Model "expected' 2020 coverage based, in the absence of COVID, based on 2000-2019 trends. 
ARIMA model uses timeseries data by country to: <br>
(a) select the most appropriate model, defined by three parameters (p, d, q) selected as follows:
  - p = number of autoregressive terms - based on minimisation of the AIC
  - d = number of non-seasonal differences needed for stationarity - based on conducting KPSS tests
  - q = number of lagged forecast errors in the prediction equation - based on minimisation of the AIC <br>
and (b) predict 2020 coverage based on this selected model

### Fit models
```{r}
    ts_df <- ts(reshape2::dcast(x, year~iso_code), start = 2000, freq = 1)
    ts_forecast <- lapply(ts_df, function(y) forecast(auto.arima(y, seasonal = FALSE), h=1, level = 95))
    
    model <- as.data.frame(do.call(rbind, ts_forecast)) 
    model=model[-1,]
    model = setDT(model, keep.rownames = "iso_code")[]
    model %<>%
      dplyr::select(iso_code, method, mean, lower, upper) 
```

Since WUENIC coverage is capped at 99% (i.e., no country can ever report 100%) we replace any 2020 estimates >99% to 99% coverage
```{r}
    model %<>%
      mutate(mean = ifelse(mean >0.99, 0.99, mean))
```


### Consolidate outputs with expected 2020 dataset
Combine forecast data with actuals to create full timeseries
```{r}
    x2 <- x %>%
      filter(year == '2000-12-31') %>%
      dplyr::select(-coverage) %>%
      mutate(year = '2020-12-31') %>%
      inner_join(model, by = "iso_code") %>%
      #dplyr::select(-c(lower, upper)) %>%
      dplyr::rename(coverage = mean, low_bound = lower, upper_bound = upper, ARIMA_model = method)
    
    x %<>%
      mutate(ARIMA_model = "WUENIC reported",
             low_bound = NA, 
             upper_bound = NA)
    
    output <- rbind(x, x2) %>%
      arrange(country, year) %>%
      mutate(coverage = as.numeric(coverage), ARIMA_model = as.character(ARIMA_model)) %>%
      mutate(plot_key = ifelse(ARIMA_model != "WUENIC reported", "Forecast", "Historic"))
    
    output %<>%
      mutate(low_bound = as.numeric(low_bound),
             upper_bound = as.numeric(upper_bound))
    
    #set-up data for t-test conducted later
    ttest_2020 <- output %>%
      filter(year == "2020-12-31") %>%
      dplyr::select(-c(ARIMA_model, country)) %>%
      dplyr::rename(predicts_2020 = coverage) %>%
      inner_join(cov_2020, by = "iso_code")
    
    #create dataset for linear models    
    cov_forecast <- ttest_2020 %>%
      as_tibble() 
    
    cov_forecast %<>%
      filter(!is.na(actuals_2020)) %>%
      dplyr::rename(predicts_lower = low_bound,
             predicts_upper = upper_bound) %>%
      mutate(cov_delta_mean = actuals_2020 - predicts_2020,
             cov_delta_high = actuals_2020 - predicts_upper,
             cov_delta_low = actuals_2020 - predicts_lower) %>%
      dplyr::select(country, iso_code, region, income_group, gavi_status, predicts_2020, 
                    predicts_lower, predicts_upper, actuals_2020, cov_delta_mean, cov_delta_high, cov_delta_low)
    
    cov_forecast %<>%
      mutate(confidence_95_level = ifelse(actuals_2020 > predicts_lower & actuals_2020 < predicts_upper,
                                   FALSE, TRUE),
             ci_width = predicts_upper - predicts_lower)
  
    write.csv(cov_forecast,"dtp3_range.csv", row.names = FALSE)
```

## Plot and visualise outputs

### For single countries
Plot single country output
```{r, message = FALSE, warning = FALSE, results = "hide", fig.keep = 1}
    country_dat <- output %>%           
      filter(country == "Jordan")       ##Replace name by country of interest
    
    country_vis = ggplot(data = country_dat, aes(x = year(year), y = coverage, color = ARIMA_model))+
      geom_point()+
      geom_smooth(colour = "darkgrey")+
      labs(title = "Jordan: WUENIC reported DTP3 coverage 2000-2019 and ARIMA-modelled expected 2020 coverage", 
           x = "Year",
           y = "Coverage (%)")+
      scale_colour_manual(name = "Datapoint:", 
                          values = c("orange", "dark grey"), 
                          labels = c("WUENIC reported coverage", "ARIMA modelled coverage"))+
      scale_y_continuous(labels = scales::percent_format(accuracy = 1L))+
      theme_bw()+
      theme(legend.position = 'bottom', 
            legend.text = element_text(size = 9),
            legend.spacing = unit(0.05, 'cm'),
            plot.title = element_text(face = "bold", size = 10),
            legend.title = element_text(face = "bold"))
    country_vis
```

### For subset of countries - largest populations
```{r, message = FALSE, warning = FALSE, results = "hide", fig.keep = 1}
    pop_dat <- output %>%
      filter(country == "China" | country == "India" | country == "Pakistan" | country == "Nigeria") %>%
      dplyr::select(country, iso_code, year, coverage, plot_key)
      #update for countries of interest

    pop_dat_2020 <- cov_2020 %>%
      filter(country == "China" | country == "India" | country == "Pakistan" | country == "Nigeria") %>%
      mutate(year = "2020-12-31") %>%
      mutate(year = as.Date(year)) %>%
      rename(coverage = actuals_2020) %>%
      mutate(plot_key = "Actuals")
    
    pop_comb <- pop_dat %>%
      rbind(pop_dat_2020) %>%
      mutate(shape = ifelse(plot_key == "Actuals", 4, 19))

    pop_vis = ggplot(data = pop_dat, aes(x = year(year), y = coverage, color = plot_key))+
      geom_point()+
      geom_point(data = pop_dat_2020, aes(x = year(year), y = coverage, color = plot_key), shape = 4)+
      geom_smooth(colour = "darkgrey")+
      labs(title = "DTP3: WUENIC reported coverage 2000-2019 and ARIMA-modelled expected 2020 coverage", 
           y = "Coverage (%)",
           x = "Year")+
      theme_bw()+
      theme(legend.position = 'bottom', 
            legend.text = element_text(size = 14),
            legend.spacing = unit(0.05, 'cm'),
            plot.title = element_text(face = "bold", size = 18),
            legend.title = element_text(face = "bold", size = 16),
            axis.text.x = element_text(angle=45,hjust = 1, size = 14),
            axis.text.y = element_text(size = 14), 
            strip.text.x = element_text(size = 14), 
            axis.title = element_text(size = 18))+
      scale_colour_manual(name = "Datapoint:", 
                          values = c("blue", "orange", "darkgrey"), 
                          labels = c("WUENIC reported 2020 coverage", "ARIMA modelled 2020 coverage", "Historic WUENIC coverage"))+
      guides(color = guide_legend(reverse=TRUE))+
      scale_y_continuous(labels = scales::percent_format(accuracy = 1L), 
                         limits = c(0.25,1),
                         n.breaks = 5)+
      facet_wrap(~country, ncol = 4, scales = "fixed", labeller = label_wrap_gen(25))
    pop_vis
    
    #save file
    ggsave(filename = "forecast_pop.png", plot = pop_vis, 
           width = 30, height = 10, units = "cm")
```

### For all countries
Output shown for 15 countries at a time - changeable by changing the paginate output
```{r, message = FALSE, warning = FALSE, results = "hide", fig.keep = 1}
    plot_dat_key <- cov_2020 %>%
      mutate(year = "2020-12-31") %>%
      mutate(year = as.Date(year)) %>%
      rename(coverage = actuals_2020) %>%
      mutate(plot_key = "Actuals")
    
    plot_dat <- output %>%
      dplyr::select(country, iso_code, year, coverage, plot_key) %>%
      rbind(plot_dat_key) %>%
      mutate(shape = ifelse(plot_key == "Actuals", 4, 19))  

    vis = ggplot(data = plot_dat, aes(x = year(year), y = coverage, color = plot_key))+
      geom_point()+
      geom_smooth(colour = "lightgrey")+
      labs(title = "DTP3: WUENIC reported coverage 2000-2019 and ARIMA-modelled expected 2020 coverage", 
           y = "Coverage (%)",
           x = "Year")+
      theme_bw()+
      theme(legend.position = 'bottom', 
            legend.text = element_text(size = 10),
            legend.spacing = unit(0.05, 'cm'),
            legend.title = element_text(face = "bold"),
            plot.title = element_text(face = "bold", size = 11),
            axis.text.x = element_text(angle=45,hjust = 1))+
      scale_colour_manual(name = "Datapoint:", 
                          values = c("blue", "orange", "darkgrey"), 
                          labels = c("WUENIC reported 2020 coverage", "ARIMA modelled 2020 coverage", "Historic WUENIC coverage"))+
      guides(color = guide_legend(reverse=TRUE))+
      scale_y_continuous(labels = scales::percent_format(accuracy = 1L), 
                         limits = c(0.3,1),
                         n.breaks = 6)+
      facet_wrap_paginate(~country, nrow = 3, ncol = 5, scales = "fixed", page = 9, labeller = label_wrap_gen(25))
    vis
    
```
<br>

### Save output as file
```{r, message = FALSE, warnings = FALSE}
    ggsave(filename = "forecast.png", plot = vis, 
           width = 25, height = 20, units = "cm")
```

## Conduct T-tests for predicted vs. observed coverage

### Remove unreliable expected coverage estimates
Countries were considered as potential unreliable datapoints based on:
1)	Visual inspection of 2000-2019 WUENIC reported coverage showed high volatility, i.e., fluctuations over a 10-percentage point range in the last five to 10 years – since this indicates coverage estimates may be unreliable and uncertain
2)	Visual inspection of forecast model fit showed fit dependent on major increases/ decreases in 2018-2019 coverage – since this indicates exceptional circumstances (e.g., stockouts, civil war) may have overly weighted forecast coverage
3)	Calculating forecast percentage point improvement in coverage from reported 2019 to forecast 2020 coverage and reviewing countries with forecast +5 percentage point improvements – since such may not be programmatically realistic (despite model fitting)

Then, WUENIC country-specific documentation was reviewed to confirm whether to remove countries

For option (3):
```{r}
    dtp3_change <- output %>% 
      filter(year == "2019-12-31" | year == "2020-12-31") %>% 
      dplyr::select(c(country, iso_code, year, coverage)) %>% 
      spread(year, coverage) %>% 
      dplyr::rename(Y2019 = "2019-12-31", Y2020 = "2020-12-31") %>% 
      mutate(delta = Y2020-Y2019)

      write.csv(dtp3_change,"Forecast_improvement_dtp3.csv", row.names = FALSE)
```


Countries considered as potential unreliable estimated coverage estimates are: <br>
(1) Criteria one - <br>
- *Solomon Islands*: high variance over time. WUENIC report that "increase in estimated coverage from 2018 [to 2019] is unlikely but exceptionally allowed given the small birth cohort size" - given such uncertainty and lack of confidence, removed
- *Suriname*: recent trends volatile, and WUENIC report "recent survey results suggest lower levels of coverage than that reported by the programme during the past 10 years. Further investigation to understand underlying differences is warranted, and WHO and UNICEF recommend a high-quality independent empirical assessment to confirm reported levels of coverage" as well as stockouts in 2019 and 2017 (one month), syringe stockout in 2018 (four months)

(2) Criteria two - <br>
- *Haiti*: Model fit skewed heavily by 2019 coverage estimate, which fell due to socio-political disturbances
- *Libya*: Model fit skewed heavily by 2018-2019 coverage estimates, which were all reduced 25 percentage points vs. country reports based on reported 3-month stockout, and therefore are not strongly tied to surveys
- *Samoa*: Model fit skewed heavily by 2018-2019 coverage estimates (much lower than historical trend). WUENIC note vaccine interruption in 2018 due public hesitancy, and that 2019 data is uncertain
- *Brazil*: estimate driven by fit to 2019 coverage (drop)which was > 15 percentage point drop in a year), and WUENIC report "no nationally representative household survey within last 5 years" so unclear how trend will evolve or if it will be confirmed

(3) Criteria three - <br>
- *Venezuela*: +6pp improvement programmatically unlikely. WUENIC coverage estimates reflect uncertainty due to civil unrest
- *El Salvador*: +6pp improvement programmatically unlikely. WUENIC report no survey in last 5 years and need for high-quality survey
- *Jordan*: +7pp improvement on 2019 coverage (from 2018-2019 there was a 10pp decrease in coverage, which was not reported as an exceptional circumstance) programmatically unlikely. 
<br>
#### Plot of removed datapoints

```{r, message = FALSE, warning = FALSE, results = "hide", fig.keep = 1}
    removed_dat <- output %>%
      filter(country == "Haiti" | country == "Libya" | country == "Samoa" | country == "Solomon Islands" 
             | country == "Brazil" | country == "El Salvador" | country == "Jordan" 
             | country == "Venezuela (Bolivarian Republic of)" | country == "Suriname")    
    
    removed_vis = ggplot(data = removed_dat, aes(x = year(year), y = coverage, color = plot_key))+
      geom_point()+
      geom_smooth(colour = "darkgrey")+
      labs(title = "DTP3: Countries removed from further analysis due to unreliable datapoints", 
           y = "Coverage (%)",
           x = "Year")+
      theme_bw()+
      theme(legend.position = 'bottom', 
            legend.text = element_text(size = 14),
            legend.spacing = unit(0.05, 'cm'),
            plot.title = element_text(face = "bold", size = 18),
            legend.title = element_text(face = "bold", size = 16),
            axis.text.x = element_text(angle=45,hjust = 1, size = 14),
            axis.text.y = element_text(size = 14), 
            strip.text.x = element_text(size = 14), 
            axis.title = element_text(size = 18))+
      scale_colour_manual(name = "Datapoint:", 
                          values = c("orange", "darkgrey"), 
                          labels = c("ARIMA modelled 2020 coverage", "Historic WUENIC coverage"))+
      theme(axis.text.x = element_text(angle=45,hjust = 1))+
      scale_y_continuous(labels = scales::percent_format(accuracy = 1L))+
      facet_wrap(~country, scales = "free", labeller = label_wrap_gen(width = 25))
    removed_vis
    
    #save file
    ggsave(filename = "removed_countries.png", plot = removed_vis, 
           width = 30, height = 30, units = "cm")
```

<br>
Remove datapoints:
```{r}
  ttest_2020 %<>%
    dplyr::filter(country != "Haiti" & country != "Libya" & country != "Samoa" & country != "Solomon Islands" & country != "Brazil"
                  & country != "El Salvador" & country != "Jordan" & country != "Venezuela (Bolivarian Republic of)" 
                  & country != "Suriname")
```


### For all countries combined
```{r}
    ttest_cov_drop <- t.test(ttest_2020$actuals_2020,
                             ttest_2020$predicts_2020,
                             paired = TRUE,
                             alternative = "two.sided")
    
    total <- ttest_2020 %>%
      filter(!is.na(actuals_2020)) %>%
      dplyr::summarize(mean(actuals_2020), mean(predicts_2020)) %>%
      mutate(vaccine = "DTP3") %>%
      dplyr::select(vaccine, "mean(actuals_2020)", "mean(predicts_2020)")
    
    total_test <- as.data.frame(do.call(cbind, ttest_cov_drop))
    total_test = total_test[-1,]
    total_test %<>%
      mutate(estimate = as.numeric(estimate), conf.int = as.numeric(conf.int), parameter = as.numeric(parameter), p.value = as.numeric(p.value)) %>%
      mutate(conf.low = (estimate + conf.int), conf.high = (estimate - conf.int))

  #Create readable output of ttest results
    tt_total <- total_test %>%
      dplyr::select(estimate, p.value, parameter, conf.low, conf.high) %>%
      dplyr::rename(point_estimate = estimate, p_value = p.value, sample_size = parameter, low_bound = conf.low, upper_bound = conf.high) %>%
      mutate(sample_size = (sample_size+1), vaccine = "DTP3") %>% #since parameter is degrees of freedom not sample size
      left_join(total, by = "vaccine") %>%
      dplyr::select(vaccine, sample_size, "mean(predicts_2020)", "mean(actuals_2020)", point_estimate, low_bound, upper_bound, p_value)

    tt_total
```
### For subset of countries where 95% confidence that coverage changed
```{r}
    confident_countries <- cov_forecast %>%
      filter(confidence_95_level == TRUE) 

    ttest_cov_drop_cleaned <- t.test(confident_countries$actuals_2020,
                             confident_countries$predicts_2020,
                             paired = TRUE,
                             alternative = "two.sided")
    
    total_cleaned <- confident_countries %>%
      filter(!is.na(actuals_2020)) %>%
      dplyr::summarize(mean(actuals_2020), mean(predicts_2020)) %>%
      mutate(vaccine = "DTP3") %>%
      dplyr::select(vaccine, "mean(actuals_2020)", "mean(predicts_2020)")
    
    total_test_clean <- as.data.frame(do.call(cbind, ttest_cov_drop_cleaned))
    total_test_clean = total_test_clean[-1,]
    total_test_clean %<>%
      mutate(estimate = as.numeric(estimate), conf.int = as.numeric(conf.int), parameter = as.numeric(parameter), p.value = as.numeric(p.value)) %>%
      mutate(conf.low = (estimate + conf.int), conf.high = (estimate - conf.int))

  #Create readable output of ttest results
    tt_total_clean <- total_test_clean %>%
      dplyr::select(estimate, p.value, parameter, conf.low, conf.high) %>%
      dplyr::rename(point_estimate = estimate, p_value = p.value, sample_size = parameter, low_bound = conf.low, upper_bound = conf.high) %>%
      mutate(sample_size = (sample_size+1), vaccine = "DTP3") %>% #since parameter is degrees of freedom not sample size
      left_join(total, by = "vaccine") %>%
      dplyr::select(vaccine, sample_size, "mean(predicts_2020)", "mean(actuals_2020)", point_estimate, low_bound, upper_bound, p_value)

    tt_total_clean
```


### By region
```{r}
    ttest_region <- ttest_2020 %>%
      filter(!is.na(actuals_2020)) %>%
      group_by(region) %>% 
      do(tidy(t.test(.$actuals_2020, 
                     .$predicts_2020, 
                     alt = "two.sided", 
                     paired = TRUE)))

    reg <- ttest_2020 %>%
      filter(!is.na(actuals_2020)) %>%
      group_by(region) %>%
      dplyr::summarise(mean(actuals_2020), mean(predicts_2020))

  #create more readable output 
    tt_reg <- ttest_region %>%
      dplyr::select(region, estimate, p.value, parameter, conf.low, conf.high) %>%
      dplyr::rename(point_estimate = estimate, p_value = p.value, sample_size = parameter, low_bound = conf.low, upper_bound = conf.high) %>%
      mutate(sample_size = (sample_size+1)) %>% #since parameter is degrees of freedom not sample size
      left_join(reg, by = "region") %>%
      dplyr::select(region, sample_size, "mean(predicts_2020)", "mean(actuals_2020)", point_estimate, low_bound, upper_bound, p_value)
    
    tt_reg
```

### By income group
```{r}
    ttest_income_group <- ttest_2020 %>%
      filter(!is.na(actuals_2020)) %>%
      filter(iso_code != "VEN") %>%
      group_by(income_group) %>% 
      do(tidy(t.test(.$actuals_2020, 
                     .$predicts_2020, 
                     alt = "two.sided", 
                     paired = TRUE)))

  #create more readable output 
    inc_grp <- ttest_2020 %>%
      filter(!is.na(actuals_2020)) %>%
      group_by(income_group) %>%
      dplyr::summarize(mean(actuals_2020), mean(predicts_2020))

    tt_inc_grp <- ttest_income_group %>%
      dplyr::select(income_group, estimate, p.value, parameter, conf.low, conf.high) %>%
      dplyr::rename(point_estimate = estimate, p_value = p.value, sample_size = parameter, low_bound = conf.low, upper_bound = conf.high) %>%
      mutate(sample_size = (sample_size+1)) %>% #since parameter is degrees of freedom not sample size
      left_join(inc_grp, by = "income_group") %>%
      dplyr::select(income_group, sample_size, "mean(predicts_2020)", "mean(actuals_2020)", point_estimate, low_bound, upper_bound, p_value)

    tt_inc_grp
```

### By Gavi status

```{r}
    tt_gavi <- ttest_2020 %>%
      filter(actuals_2020 >= 0) %>%
      group_by(gavi_status) %>% 
      do(tidy(t.test(.$actuals_2020, 
                     .$predicts_2020, 
                     alt = "two.sided", 
                     paired = TRUE)))
    
    gav <- ttest_2020 %>%
      filter(actuals_2020 >= 0) %>%
      group_by(gavi_status) %>%
      dplyr::summarize(mean(actuals_2020), mean(predicts_2020))
    
  #create more readable output 
    tt_gavi_grp <- tt_gavi %>%
      dplyr::select(gavi_status, estimate, p.value, parameter, conf.low, conf.high) %>%
      dplyr::rename(point_estimate = estimate, p_value = p.value, sample_size = parameter, low_bound = conf.low, upper_bound = conf.high) %>%
      mutate(sample_size = (sample_size+1)) %>% #since parameter is degrees of freedom not sample size
      left_join(gav, by = "gavi_status") %>%
      dplyr::select(gavi_status, sample_size, "mean(predicts_2020)", "mean(actuals_2020)", point_estimate, low_bound, upper_bound, p_value)
    
    tt_gavi_grp

```

## Interpretation outputs
### Tables for greatest declines in coverage
```{r}
    cleaned_forecast_dtp3 <- ttest_2020 %>%
      dplyr::select(c(country, iso_code, region, income_group, gavi_status, predicts_2020, actuals_2020)) %>%
      mutate(cov_delta = actuals_2020 - predicts_2020) %>%
      mutate(income_group = dplyr::recode_factor(income_group, 
                                                  "Low income" = "LIC", 
                                                  "Lower middle income" = "LMIC", 
                                                  "Upper middle income" = "UMIC", 
                                                  "High income" = "HIC")) %>%
      mutate(region = factor(region,
                             levels = c("Western Europe",
                                        "Eastern Europe & Central Asia",
                                        "Eastern & Southern Africa",
                                        "Middle East & North Africa",
                                        "West & Central Africa",
                                        "South Asia",
                                        "East Asia & Pacific",
                                        "Latin America & Caribbean",
                                        "North America"))) %>%
      mutate(gavi_status = dplyr::recode_factor(gavi_status,
                                                "Initial self-financing"= "Gavi - ISF",
                                                "Preparatory transition phase" = "Gavi - PT",
                                                "Accelerated transition phase" = "Gavi - AT",
                                                "Fully self-financing" = "Gavi - FSF",
                                                "Never Gavi" = "Never Gavi")) %>%
      filter(!is.na(actuals_2020))
      
      cov_change_tab <- cleaned_forecast_dtp3 %>%
        mutate(actuals_2020 = percent(actuals_2020, accuracy = 0.1),
               predicts_2020 = percent(predicts_2020, accuracy = 0.1),
               cov_delta = percent(cov_delta,accuracy = 0.1))
```

### Summary modelled impact on RI
- Import UNWPP data
- Multiply coverage delta estimates by surviving infant populations to calculate additional missed immunisations
- Create summary output table including % coverage delta and  missed immunisations 
```{r}
file_path_unwpp <- here::here("data/unwpp.csv")

unwpp <- file_path_unwpp %>%
  rio::import(header = TRUE) %>%
  tibble()
unwpp

unwpp %<>%
  mutate(iso_code = countrycode(sourcevar = Country, origin = "country.name", destination = "iso3c")) %>%
  filter(!is.na(iso_code)) %>%
  dplyr::select(-Country)

cov_forecast %<>%
  left_join(unwpp, by = "iso_code") %>%
  dplyr::rename(surviving_infants = "Surviving infants") %>%
  mutate(additional_missed = as.integer(surviving_infants*cov_delta_mean*-1),
         expected_missed = as.integer((1-predicts_2020)*surviving_infants),
         total_missed = as.integer((1-actuals_2020)*surviving_infants),
         unreliable_flag = ifelse(country %in% removed_dat$country, TRUE, FALSE))

appendix_output <- cov_forecast %>%
  mutate(actuals_2020 = percent(actuals_2020, accuracy = 0.1),
         predicts_2020 = percent(predicts_2020, accuracy = 0.1),
         cov_delta_mean = percent(cov_delta_mean,accuracy = 0.1)) %>%
  rename(Country = country,
         "ISO code" = iso_code,
         Region = region,
         "Income group" = income_group,
         "Gavi eligibility status" = gavi_status,
         "ARIMA modelled expected 2020 coverage" = predicts_2020,
         "WUENIC reported 2020 coverage" = actuals_2020,
         "Coverage delta" = cov_delta_mean,
         "Surviving infants" = surviving_infants,
         "Additional missed children" = additional_missed,
         "Expected missed children" = expected_missed,
         "Total missed children" = total_missed,
         "Data unreliable flag" = unreliable_flag)

write.csv(appendix_output,"SupplementaryFiles_DTP3_coverage_impact.csv", row.names = FALSE)

```

### Summary modelling output for Appendix
```{r}
 order <- output %>%
  filter(plot_key == "Forecast") %>%
  dplyr::select(country, iso_code) %>%
  left_join(model, by = "iso_code") %>%
  #inner_join(cov_2020, by = c("iso_code", "country")) %>%
  as_tibble() %>%
  mutate(mean = percent(as.numeric(mean), accuracy = 0.1),
         lower = percent(as.numeric(lower), accuracy = 0.1),
         upper = percent(as.numeric(upper),accuracy = 0.1),
         #actuals_2020 = percent(as.numeric(actuals_2020),accuracy = 0.1),
         #cov_delta = percent((as.numeric(actuals_2020) - as.numeric(mean)), accuracy = 0.1)
         )%>%
  rename(Country = country,
         "ISO code" = iso_code,
         "ARIMA model fit" = method,
         "Expected 2020 coverage" = mean,
         "Expected: 95% Confidence interval - lower bound" = lower,
         "Expected: 95% Confidence interval - upper bound" = upper,
         #"WUENIC 2020 reported coverage" = actuals_2020,
         #"Coverage delta" = cov_delta
         )

order <- apply(order,2,as.character)

write.csv(order,"SupplementaryFiles_DTP3_coverage_model.csv", row.names = FALSE)
```

### Save files
```{r}
    write.csv(cov_forecast,"forecast_dtp3.csv", row.names = FALSE)
    write.csv(tt_reg,"ttest_region_dtp3.csv", row.names = FALSE)
    write.csv(tt_inc_grp,"ttest_income_dtp3.csv", row.names = FALSE)
    write.csv(tt_gavi_grp,"ttest_gavi_dtp3.csv", row.names = FALSE)
    write.csv(tt_total,"ttest_total_dtp3.csv", row.names = FALSE)
    write.csv(tt_total_clean,"ttest_total_dtp3_clean.csv", row.names = FALSE)
    write.csv(cleaned_forecast_dtp3, "cleaned_forecast_dtp3.csv", row.names = FALSE)
    write.csv(cov_change_tab,"paper_coverage_table_dtp3.csv", row.names = FALSE)
```

